\documentclass[twocolumn]{aastex631}
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\bibliographystyle{aasjournal}
\usepackage{amsmath}
\usepackage{graphicx}
% \usepackage{caption}
\input{figures.tex}
\begin{document}
\title{Diffusion Based 3D Reconstruction of the Cosmic Baryons}

\author[0000-0001-9580-1043]{Maryam Hussaini}
\email{maryam.hussaini@cfa.harvard.edu} % from Liam: only list your email, you are corresponding author.
\affiliation{Center for Astrophysics $\mid$ Harvard \& Smithsonian, Cambridge, MA 02138, USA}

\author[0009-0004-3662-2263]{Aizhan Akhmetzhanova}
%\email{liam.connor@cfa.harvard.edu}
\affiliation{Center for Astrophysics $\mid$ Harvard \& Smithsonian, Cambridge, MA 02138, USA}

\author[0000-0002-6069-2999]{Carolina Cuesta-Lazaro}
%\email{liam.connor@cfa.harvard.edu}
\affiliation{Center for Astrophysics $\mid$ Harvard \& Smithsonian, Cambridge, MA 02138, USA}

\author[0000-0002-7587-6352]{Liam Connor}
%\email{liam.connor@cfa.harvard.edu}
\affiliation{Center for Astrophysics $\mid$ Harvard \& Smithsonian, Cambridge, MA 02138, USA}


\begin{abstract}
The complex relationship between galaxies and diffuse gas at large scales depends strongly on astrophysical feedback and cosmology. The dispersion of Fast Radio Bursts (FRBs) is a column of cosmic baryon density and is dominated by the intergalactic medium, which is not necessarily captured by traditional halo model prescriptions. We introduce a diffusion model to reconstruct 3D gas density fields from observed galaxy surveys, including redshift space distortions. The probabilistic model is trained on the CAMELS50 simulation suite (IllustrisTNG) and learns to predict the posterior distribution of gas fields conditioned on stellar mass-weighted galaxy density, cosmological parameters, and the baryon suppression powerspectrum. We validate the model on test boxes by showing agreement between predicted and true dispersion measure across individual lines of sight and the overall dispersion measure distribution.
The model generalizes to a box size $\sim\,4^3$ times larger than the training boxes and can be used to reconstruct the line of sight to cosmological FRBs. The model's 3D gas density predictions vary with input power spectrum and feedback scenarios, opening a new inference path to constrain feedback models from FRBs and other probes of the cosmic baryons. 
\end{abstract}

\section{Introduction}
\label{sec:intro}

The distribution of cosmic baryons has remained a pressing problem in extragalactic astronomy for decades, due to a lack of 
direct measurements and its role in our understanding of galaxy formation and precision cosmology. The majority 
of this gas resides in a diffuse ionized state in the intergalactic medium (IGM) and in dark matter halos. 
The precise distribution of the gas is modulated 
by complex astrophysical processes in galaxy evolution. In recent years, new probes of the cosmic baryons have begun 
to shed light on gas at the scales of halos and above \citep{bigwood2024, popesso2026, hadzhiyska2024, FRBs}. Several measurements 
suggest that strong AGN feedback depletes the 
baryon budgets of group-scale halos \citep{XX, mccarty2026}.

Supernovae, AGN, and \textbf{cosmic rays} inject energy into 
halo gas, heating it, creating winds, and pushing the gas out of galaxies and halos, or preventing its accretion and cooling. \textbf{These processes regulate star formation and the growth of supermassive black holes, leading to non-linear galaxy feedback that remains poorly constrained.} 
%It is the regulating mechanism that keeps galaxies from growing too much and forming too many stars. 
In simulations, the choice of inaccurate feedback
prescriptions result in discrepancy with 
observations in the stellar mass function, halo gas fractions, baryon density profiles, and temperature distributions. Feedback sits at the intersection of galaxy formation and precision cosmology. In galaxy formation feedback regulates how baryons turn into stars, the nature of the CGM, and determines galaxy populations. In cosmology, feedback is the contaminant that breaks the coupling between dark matter and baryons. Feedback redistributes baryons and changes the baryonic power spectrum from that of the dark matter via suppression of power at small scales. Understanding feedback offers a path to answering questions in galaxy formation and alleviates a key systematic in weak lensing cosmology surveys.

Given that feedback and cosmology are responsible for 
the distribution of cosmic baryons, methods that allow us to infer the 3D distribution of gas are valuable for 
myriad areas in extragalactic astronomy [cite cite cite]. Reconstruction of the dark matter field and velocity field from galaxy surveys has made significant advances in the past decade \citep{hong2021, Villanueva2021, mudur2022, park2023}[cite for velocity field cite cite]. The 3D reconstruction of the cosmic baryons is comparatively unexplored. \textbf{Liam comment: add a couple of sentences on what's been done so far on baryon reconstruction, but don't mention FRBs yet.}


One recent, incisive probe of the cosmic baryon distribution is the dispersion of Fast Radio Bursts (FRBs). 
%To connect a reconstructed gas field to observations, we use fast radio bursts (FRBs), a powerful probe of ionized baryons. 
FRBs are bright, millisecond-duration radio transients originating from galaxies across the Universe \citep{lorimer2007, petroffreview}. As the burst propagates toward us, its radio signal interacts with the ionized gas it encounters along the way and higher frequencies arrive earlier than lower frequencies, an effect known as dispersion. The magnitude of this delay, the Dispersion Measure (DM), is directly proportional to the integrated free electron column density along the line of sight. For a source at redshift $z_s$, the mean cosmic contribution is
\begin{equation}
    \langle \mathrm{DM_{cosmic}}(z_s) \rangle = \int_0^{z_s} \frac{c \, \bar{n}_e(z) \, (1+z)}{H(z)} \, dz,
\end{equation}
where $\bar{n}_e(z)$ is the mean free electron density, $H(z)$ is the Hubble parameter, and $c$ is the speed of light. Because $\bar{n}_e$ is set by the cosmic baryon density, the DM-redshift relation directly traces the integrated ionized baryon content of the Universe \citep{macquart2020}. In its journey to Earth, an FRB signal accumulates DM contributions from the gas in its host galaxy, the intergalactic medium, the circumgalactic medium of intervening halos, and the Milky Way. However, the DM alone tells us only the total amount of gas along a sightline and not the 3D distribution of that gas, which is precisely what encodes the feedback signal we want.


Mapping from galaxies to gas is inherently uncertain because a given galaxy field can have a variety of gas configurations. So, beyond producing a single reconstruction, we would like a method that captures the uncertainty by producing a posterior distribution of gas rather than a single estimate with no error bars. Probabilistic models, and generative diffusion models in particular, have dramatically improved our ability to reconstruct 3D cosmic fields from galaxy observations. In this work, we adapt and optimize an existing diffusion-based framework, previously used to reconstruct the dark matter field \citep{ono2024}, for 3D gas field reconstruction tailored to FRB science. We train the model on the CAMELS simulation suite, which spans a wide range of cosmological and feedback scenarios and therefore lets the model marginalize over the uncertainties associated with the feedback. The reconstructed gas fields return a full posterior, and combined with FRB dispersion measures, they allow us to constrain where the baryons are and how they got there. Recent efforts have begun this program along individual FRB sightlines: the FLIMFLAM survey used spectroscopic data from foreground galaxies to constrain the partition of baryons between the IGM and the CGM of intervening halos \citep{Khrykin_2024, kglee23, simha2020, simha2023}.


Beyond galaxies, we condition our model on the baryon suppression spectrum $sP(k) = P_{\text{hydro}}/P_{\text{DM}}$. this ratio captures how feedback has suppressed the matter power spectrum compared to the dark matter only spectrum. $sP(k)$ is the scale resolved quantity that gas distribution responds to \citep{sharma2025}. Conditioning on $sP(k)$ lets the model discriminate between feedback scenarios and produce different gas maps depending on the provided $sP(k)$. 
% \textbf{The next few paragraphs' outline}
% \begin{itemize}
%     \item FRBs probe ionized baryons.The DM equation. 3D gas reconstruction lets us predict and interpret FRB sightlines 
%     \item diffusion model returns a full posterior. we use CAMELS \cite{camels2021}. Camels lets us marginalize over feedback uncertainty (holy shit this is so important and cool). This builds on diffusion reconstruction work done by \cite{ono2024, mudur2022, park2023}
%     \item we condition on the baryon suppression spectrum so the model can discriminate feedback scenarios \cite{sharma2025}
% \end{itemize}
% a powerful


% Furthermore, we want a method that quantifies uncertainty, too. %diffusion model enters the chat. 


% % Why is it useful to have a 3D reconstruction algorithm? 
% Most of the baryonic matter in the Universe is thought to reside in the intergalactic medium (IGM) and circumgalactic medium (CGM). This diffuse gas is heated, blown out, and redistributed by stellar and AGN feedback processes. Because it is so diffuse, it is extremely difficult to observe directly, and as a result, the 3D distribution of cosmic baryons remains poorly constrained. A better map of this gas would deepen our understanding of how feedback operates across cosmic time, which is an open questions in galaxy formation. It would also allow us to better interpret the dispersion measures of Fast Radio Bursts, and determine what fraction of each FRB's DM originates from the IGM, the CGM of intervening halos, or the host galaxy.

% At the same time, we are entering an era of unprecedented galaxy survey data. The Dark Energy Spectroscopic Instrument (DESI) has completed its 5-year survey, obtaining spectroscopic redshifts of 50 million galaxies\citep{desi}. Upcoming surveys like Rubin/LSST will detect billions more. Galaxies trace the underlying matter distribution on large scales, but their relationship to the diffuse cosmic baryons is less well understood because this gas is difficult to observe. \citet{lee2022} demonstrated that foreground data increases the sensitivity of localized FRBs to cosmological and astrophysical parameters by $\sim\times25$. A reconstruction algorithm that maps galaxy surveys directly to the 3D gas distribution would unlock the scientific potential of this data, enabling us to rule out feedback models, interpret FRB sightlines, and map the cosmic baryon distribution.

% Fast Radio Bursts (FRBs) are bright, millisecond long radio transients originating from galaxies across the Universe \citep{lorimer2007, petroffreview}. As the burst propagates toward us, its radio signal interacts with the ionized gas it encounters along the way and higher frequencies arrive earlier than lower frequencies, an effect known as dispersion. The magnitude of this delay, the Dispersion Measure (DM), is directly proportional to the integrated free electron column density along the line of sight. In its journey to Earth, an FRB signal accumulates DM contributions from the gas in its host galaxy, the intergalactic medium, the circumgalactic medium of intervening halos, and the Milky Way. FRBs are therefore a uniquely powerful probe of diffuse ionized gas that is otherwise extremely difficult to detect \citep{macquart2020}. However, the DM alone tells us only the total amount of gas along a given sightline and not where that gas came from. Knowledge of the 3D gas distribution along the line of sight can help us distinguish the DM contributions of the IGM, the CGM of intervening halos, and the host galaxy.  

% Recent advances in machine learning have dramatically improved our ability to reconstruct 3D cosmic fields from galaxy observations. Combined with  
% FRB dispersion measures, these reconstructions allow us to constrain where the baryons are and how they got there, helping us set limits on cosmological and astrophysical parameters (\citep{Khrykin_2024,kglee23,simha2020,simha2023}). 
% In this work, we adapt and optimize an existing diffusion based reconstruction framework for 3D gas field reconstruction, tailored  for FRB science. By conditioning the model on the baryon suppression spectrum   we take a step toward using gas reconstructions to discriminate between feedback models, opening a new path for connecting FRB observations to the physics of galaxy formation.

In this work, we develop a diffusion model that reconstructs 3D gas fields from galaxies, conditioned on cosmology and the suppression spectrum, returning a full posterior; we validate it on CAMELS, show it generalizes to TNG300, and demonstrate FRB DM prediction. This paper is organized as follows: Section~\ref{sec:methods} describes the simulation data, preprocessing pipeline, and model architecture; Section~\ref{sec:results} presents validation results on held-out CAMELS50 test boxes and generalization to TNG300; Section~\ref{sec:discussion} discusses implications for FRB science and limitations; and Section~\ref{sec:conclusion} summarizes our conclusions and future directions.

% % What is an FRB and why do astronomers care about it?
% FRBs are bright, short-duration (ms), transient, radio signals coming from all over the universe from other galaxies. The radio signal of the burst interacts with the ionized diffuse gas it encounters on its path, and higher frequencies arrive later than lower frequencies. The measurement of this delay, called the dispersion measure, can tell us how much ionized gas was along the line of sight of the FRB. In its path, the FRB interacts with gas in its host galaxy, the intergalactic medium, circumgalactic medium of other galaxies it passes by, and the gas in the Milky Way. By measuring the dispersion measure, we can determine the total sum of all these matters. This is important because some of the matter in the universe, like the intergalactic medium, is very hard to detect by other methods given that it is diffuse and hot. But with FRBs, we can prove this diffuse matter and measure the quantity of it. 

% % Why do we need to know what's in the gas between galaxies?

% Most of the matter in the universe is between galaxies, in the intergalactic medium. Having a reconstruction of the gas between galaxies is mapping the "missing baryons." Understanding how that matter is structured will help us understand how supernova and AGN feedback work. Having a good reconstruction of the gas between galaxies will help us discriminate against unphysical feedback models. 

% % How do we currently try to reconstruct the gas between galaxies?
% FRBs are bright, short-duration (ms), transient, radio signals coming from all over the universe from other galaxies. The radio signal of the burst interacts with the ionized diffuse gas it encounters on its path, and higher frequencies arrive later than lower frequencies. The measurement of this delay, called the dispersion measure, can tell us how much ionized gas was along the line of sight of the FRB. In its path, the FRB interacts with gas in its host galaxy, the intergalactic medium, circumgalactic medium of other galaxies it passes by, and the gas in the Milky Way. By measuring the dispersion measure, we can determine the total sum of all these matters. This is important because some of the matter in the universe, like the intergalactic medium, is very hard to detect by other methods given that it is diffuse and hot. But with FRBs, we can prove this diffuse matter and measure the quantity of it. 3D reconstruction of the cosmic web is so helpful when using FRBs as probes for the large scale structure. Luckily, due to advances in machine learning, the models for 3D reconstruction have gotten so much better. We are optimizing an existing algorithm to use it for 3D gas reconstruction and tailoring it to FRB studies. By training the model in a feedback conditional way, we will one day be able to use reconstruction of the gas to discriminate against unfavorable feedback parameters. 
 

% To test if our model could be generalized to bigger box sizes, we ran inference on TNG boxes. Our model had trained on the TNG suite in the CAMELS simulation, so it should work equally as well on a TNG input. The box size that our model trained on was 50 mpc $^3$, but we ran inference on boxes as large as 302 mpc$^3$ and recovered good reconstructions. 
% \\
% The dispersion measure (DM) is a measure of the total amount of gas along the line of sight toward a source. It is defined as the integral of electron density along the path:
% \begin{equation}
% \mathrm{DM} = \int_0^d n_e(l)\,dl
% \end{equation}
% where $n_e(l)$ is the electron density at distance $l$ along the line of sight. In practice, DM tells us how much ionized gas the radio signal passed through. In our simulations, we compute DM by summing the gas density along one spatial direction (the line of sight), which produces a 2D map where each pixel value represents the total gas column in that direction. This makes DM a key observable: FRBs measure DM directly, and our reconstructed gas fields allow us to predict what DM values should be for any location in the universe.
\figOverview
\figInput




\figCigar


\section{Methods}
\label{sec:methods}

We train a conditional flow matching model to reconstruct the 3D gas density field from the observed galaxy density field, conditioned on cosmological parameters and the baryon suppression spectrum. We use the CAMELS simulation suite (IllustrisTNG), which provides 1000 simulation boxes each with a volume of $(50\,h^{-1}\,\mathrm{Mpc})^3$ discretized into $128^3$ voxels. Each box contains the 3D stellar mass-weighted galaxy density field, the 3D gas density field, and 36 cosmological and astronomical parameters.

We preprocess the galaxy field by applying a minimum stellar mass threshold $M_{\star} < 10^9\,M_\odot$, populating a $128^3$ voxel grid using Cloud-In-Cell (CIC) interpolation weighted by $\log_{10}(M_{\star})$, computing $\log_{10}$ of the resulting density field, and normalizing by the CAMELS50 training statistics: $\tilde{f}_{\mathrm{gal}} = (\log_{10}(f_{\mathrm{gal}}) - \mu)/\sigma$, where $\mu = -5.94$ and $\sigma = 0.586$.

The gas density field is preprocessed by first applying a $\log_{10}$ transformation (with a small offset of $10^{-6}$ to avoid $\log(0)$), then standardizing to zero mean and unit variance using statistics computed from the training data ($\mu = 9.5$, $\sigma = 0.56$ for the $128^3$ resolution).


For each training box, we adopt the baryon suppression spectrum $sP(k) = P_\mathrm{hydro}(k)/P_\mathrm{DM}(k)$, the ratio of the matter power spectrum in the full hydrodynamical simulation to that of its gravity only ($N$-body) counterpart, computed with \texttt{Pylians}\citep{villaescusa-navarro2018}. This ratio quantifies the scale-dependent redistribution of matter by baryonic physics and feedback, which suppresses power on scales $k \sim 1 - 50\,h\,\mathrm{Mpc}^{-1}$, and enhances in smaller, galactic scales\citep{genel2026}. It encodes the combined signature of cosmological parameters (which set the baryon abundance) and feedback strength (which sets the redistribution), providing a more sensitive conditioning signal than individual feedback parameters alone.

We use a conditional flow matching model to learn the posterior $p_\theta(n_e(\mathbf{x}) \mid \delta_g(\mathbf{x}), \mathbf{\Omega}, sP(k))$. Flow matching learns to transport samples from a noise distribution $x_0 \sim \mathcal{N}(0, \mathbf{I})$ to the target data distribution $x_1$ along a linear interpolation path $x_t = (1-t)x_0 + t x_1$, $t \in [0,1]$, by minimizing the mean-squared error between the predicted and true velocity fields $u_t = x_1 - x_0$. At inference time, the learned velocity field is integrated using a neural ODE solver with 50 Euler steps. The backbone is a 3D conditional U-Net operating on $128^3$ voxel grids, with three encoder/decoder levels at channel dimensions $[32, 64, 64]$, $3\times3\times3$ circular-padded convolutions, group normalization, and SiLU activations. The galaxy field is concatenated channel-wise with the target field; cosmological parameters and $sP(k)$ are embedded via sinusoidal positional encodings and broadcast to all U-Net blocks. The model is trained with AdamW at a learning rate of $10^{-4}$, effective batch size 16 (4 gradient accumulation steps), $\ell_2$ gradient clipping at 1.0, and 16-bit mixed precision for 30,000 gradient steps. We use 980 boxes for training, 10 for validation, and 10 for testing.

\textbf{Describe how we go from ne to DM. Describe then how we compute p(DM) from p(ne). Say that we can 
generate DM maps from galaxy positions or different cosmological parameters, 
power spectra. Use equations.}
The dispersion measure quantifies the total column of free electrons along the line of sight toward a source, and is defined as

\begin{equation}
\mathrm{DM} = \int_0^d \frac{n_e(l)}{1+z}\,dl,
\end{equation}
where $n_e(l)$ is the electron density at distance $l$ along the path and the factor $(1+z)$ accounts for cosmological redshifting of the dispersion. For our reconstructed fields, we convert the gas density to a free electron number density and integrate along one spatial axis of the box (the line of sight direction), accounting for the physical voxel path length. This collapses the 3D field into a 2D map in which each pixel gives the total electron column, in $\mathrm{pc\,cm^{-3}}$, in that direction.
% Because DM is a linear functional of $n_e$, the posterior over gas fields produced by our model maps directly to a posterior over DM. Each reconstruction is a sample $n_e(\mathbf{x})$ where $\mathbf{x}$ is a 3D position vector; integrating it along the line of sight yields one DM value, so a set of reconstructions gives a set of DM values that are samples from the DM posterior. No additional modeling is required to propagate the uncertainty, since the line of sight integral is a simple weighted sum.

Our model produces a posterior over gas fields conditioned on the observed galaxy field, cosmology, and the suppression spectrum:

\begin{equation}
n_e(\mathbf{x}_\perp, l) \sim p_\theta\!\left(n_e \mid \delta_g, \boldsymbol{\Omega}, sP(k)\right),
\end{equation}

where $\mathbf{x}_\perp$ denotes the two sky coordinates and $l$ the line-of-sight coordinate. The dispersion measure at each sky position is the line-of-sight integral of the electron density,

\begin{equation}
\mathrm{DM}(\mathbf{x}_\perp) = \int_0^d \frac{n_e(\mathbf{x}_\perp, l)}{1+z}\, dl.
\end{equation}

Because this integral is a linear functional of $n_e$, applying it to each posterior sample yields a corresponding sample of DM, so the gas-field posterior induces a posterior over dispersion measure directly:

\begin{equation}
\mathrm{DM}(\mathbf{x}_\perp) \sim p_\theta\!\left(\mathrm{DM} \mid \delta_g, \boldsymbol{\Omega}, sP(k)\right).
\end{equation}

No additional modeling is needed to propagate the uncertainty, since DM is obtained from $n_e$ by a simple weighted sum along the line of sight.

To test generalization, we apply the trained model to TNG300, a simulation four times larger than the training boxes in volume $(302\,h^{-1}\,\mathrm{Mpc})^3$. We preprocess the TNG300 galaxy field using the same pipeline as above, rescaling galaxy positions by a factor of $50/35$ to match the CAMELS50 voxel resolution, and apply the same normalization statistics from training.

All simulations are evaluated at redshift $z = 0$. Dispersion measures are computed by summing the reconstructed gas density field along the line of sight, consistent with the definition of DM as the integral of electron column density.

\figPatchHist

\section{Results}
\label{sec:results}
Figure~\ref{fig:input_output} shows the model's reconstruction quality on a CAMELS50 test box. Although the model reconstructs 3D gas density, 
the observable of DM is a collapsed quantity that integrates in the FRB direction. We sum the reconstructed gas along the line of sight to produce 2D dispersion measure maps. We find 
that our reconstructed DM map recovers the large scale structure of filaments, nodes, and voids present in the True DM map, even with sparse galaxy inputs. To quantify this, the bottom panels of Figure~\ref{fig:input_output} show a scatter plot of the posterior mean DM against the true DM across all lines of sight in the box, and a histogram comparing the full DM distributions. The scatter is tightly clustered around the 1:1 line with $\sigma = 0.124$, and the reconstructed DM distribution closely matches the true distribution. The model accurately reconstructs both individual sightlines and the overall statistics of the gas field.
\figRank

A key benefit of probabilistic models like ours is that they produce a posterior distribution, not just a mean reconstruction. DM maps from the mean reconstruction tend to wash out small-scale structure, because in patches with low input galaxy density the model is uncertain and the mean averages over many possibilities. Individual samples from the reconstruction posterior are more visually similar to the true gas, because the learned prior demands the narrow filaments and low-density halos it has seen during training. To validate the model's probabilistic predictions, Figure~\ref{fig:patch_hist_box9} shows 150 independent samples drawn from the model at three representative locations: a void, a filament, and a dense node. In all three environments, the model's samples cluster around the ground truth value. Because the model provides a distribution of voxel values rather than a point estimate, uncertainty quantification on FRB DMs follows naturally. We find that the true DM value is consistently positioned within the predicted distribution, demonstrating that the model has learned realistic uncertainties: the posterior is neither overconfident (a sharp peak that could miss ground truth) nor uninformative (spanning an unreasonably wide range).
\\
Next, we assess the model's ability to preserve the relative ordering of DM values. We computed the rank correlation between true and reconstructed dispersion measures. Figure \ref{fig:rank} shows a 2D histogram comparing the rank of true DM values to the rank of reconstructed DM values across all lines of sight in a test box. High density concentrated along the diagonal (white dashed line) indicates that the model accurately preserves the relative ordering of DM values from lowest to highest. This demonstrates that even where absolute DM predictions may deviate from the truth, the model correctly identifies which regions that or over- or under-dense. This rank based metric confirms that the model captures the amplitude distribution and relative structure of the cosmic web.\\

\figFeedback

Figure~\ref{fig:pk} explores how the model responds to different feedback scenarios. We fed the same galaxy field into the model three times, each with a different suppression spectrum: high, correct, and low feedback. With high feedback, the model produces a smoother, more diffuse gas, as expected when strong feedback has blown gas out of dense structures into the voids. With the correct spectrum matching the test box, the reconstruction matches the true gas distribution. With low feedback, the model produces a more concentrated distribution with tighter filaments and denser nodes. 

\subsection{Scaling to TNG300}
To test whether the model generalizes beyond the training distribution, we applied it to TNG300, a simulation box four times larger than the CAMELS50 training data (205 vs 50 $h^{-1}$ Mpc on a side). We rescaled the TNG300 data using the same preprocessing pipeline as the training data. \textbf{Our diffusion model is convolutional and therefore does not require significant changes for different 
input box size.} Figure \ref{fig:tng300} shows that the model successfully reconstructs the DM distribution across the larger box. The reconstructed DM histogram closely matches the true distribution, which demonstrates that the model learned generalizable physics rather than memorizing training patterns, and can scale to significantly larger volumes than those used for training. Although the model shows a systematic underprediction of approximately 0.1 in $\log_{10}(\mathrm{DM})$ on average which could be due to the normalization process (we normalize using CAMELS50 mean/std, but TNG300 might have different gas statistics. Ask Aizhan.)
As a way to show how the model can be used to study individual FRB lines of sight, we stitched together 10 CAMELS50 boxes to creat a line of sight that extends to $\sim500\,h^{-1}\,\mathrm{Mpc}$. Figure \ref{fig:cigar} shows the reconstructed vs the true gas of the test boxes. So, the model can be generalized to large boxes like we showed in \ref{fig:tng300}, but we can also stitch multiple boxes together to create a long line of sight and that would still be informative. 

\figTNG

% \figModelComparison
\section{Discussion}
\label{sec:discussion}

In generalization of our model to TNG 300, and also with the CAMELS test boxes, the distribution of dispersion measure in the reconstruction has a systematic offset from that of the true distribution (Figure \ref{fig:tng300}). At the same time, the rank correlation in Figure \ref{fig:rank} demonstrates that the relative ordering of the DM values is accurate with a Spearman rank correlation of $\rho = 0.8850$. This validates that the relative gas distribution in the reconstruction is trustworthy. Therefore, for the FRB science applications, this is not problematic because given the known value of $\Omega_b$ from cosmology and observed DM of an FRB, we can calibrate out the offset and extract accurate absolute DM. 

Our efforts to reconstruct FRB sightlines in hopes of ruling out feedback models and constraining cosmological and astrophysical parameters are in line with existing surveys such as FLIMFLAM \citep{Khrykin_2024} which constrained the IGM and CGM baryon fractions along 8 FRB sightlines by taking the spectra of all the foreground galaxies. With our reconstructions, all the subsequent science of constraining the parameters and partitioning DM contributions between the IGM and CGM remains the same.  Given the fact that we only need photometric galaxy data to produce the reconstructions, which makes our model scalable to a larger number of FRBs, we will have a large enough sample size to rule out feedback models inconsistent with observations. Although CAMELS parameters span a broad range by design, the inferred parameters are still limited to the simulation's prior range, which is a caveat of all simulation based inference models. 

\textbf{Liam:} Our 
probabilistic 3D reconstruction algorithm has been 
verified on simulated data under a series of conditions. We 
have not yet tested the method on real galaxy 
survey data, opting to leave that to a future effort in which 
observational selection effects can be carefully considered. 
Our current training and validation data assume uniform 
galaxy sampling and redshift-independent thresholds on stellar mass. While our galaxy number densities are comparable to 
existing spectroscopic galaxy surveys, we expect model 
mismatch between training and validation data to degrade 
reconstruction quality unless training data is forward modeled 
to match specific surveys. 

Limitations: \\
1. the minimum stellar mass threshold used to define the galaxy sample is not calibrated to a specific observational survey. The reconstruction is sensitive to this choice as shown in figure xxx.\\
2. the model has the same systematic assumptions and biases present in the CAMELS IllustrisTNG simulations. If the real Universe differs from these assumptions, the reconstructions will be biased. \\
3. The model is conditioned on cosmological parameters assumed to be known exactly, whereas in practice they carry observational uncertainties. \\



\section{Conclusions}
\label{sec:conclusion}
We have introduced a diffusion-based model for reconstructing the 3D distribution of cosmic baryons from galaxy survey data, conditioned on cosmological parameters and the baryon suppression spectrum. Trained on the CAMELS50 simulation suite, the model learns to predict the posterior distribution of gas density fields given stellar mass-weighted galaxy density. We validated the model on test boxes, demonstrating accurate reconstruction of both the spatial structure and statistical distribution of the gas field, as measured by dispersion measure (DM) along lines of sight. The model generalizes to larger box sizes, successfully reconstructing DM distributions in TNG300. By conditioning on the baryon suppression spectrum, the model captures the integrated effects of feedback processes, allowing for future applications in constraining feedback models from FRB observations. This framework opens a new avenue for mapping the cosmic baryons and leveraging FRBs as probes of large-scale structure and galaxy evolution.

Future work will explore the sensitivity of reconstructions to different galaxy selection functions, training on multiple simulation suites to mitigate model dependence, and incorporating cosmological parameter uncertainties into the conditioning framework. Despite these caveats, our results demonstrate that diffusion models can effectively reconstruct the 3D distribution of cosmic baryons from galaxy observations, providing a powerful new tool for interpreting FRB measurements and probing the cosmic web.

\section{ACKNOWLEDGMENTS}

\clearpage

\bibliography{ref}
\end{document}